Missing the Point\(\colon\) Non-Convergence in Iterative Imputation Procedures

### 2020-04-09 3.Thesis/1.SimulationStudy/Writeup/manuscript.Rmd

<!-- The number of imputations is denoted by $m$ (i.e., the number of times each datapoint in $y_{mis}$ is filled in). The aim of MI is to impute each datapoint in $y_{mis}$ $m$ times. Because of the iterative nature of the MICE algorithm -->

>>>>

**Add: missingness percentage = proportion of cases with one or more missing values times 100 **
<!-- The number of imputations is denoted by $m$ (i.e., the number of times each datapoint in $y_{mis}$ is filled in). The aim of MI is to impute each datapoint in $y_{mis}$ $m$ times. Because of the iterative nature of the MICE algorithm -->

### 2020-04-09 3.Thesis/1.SimulationStudy/Writeup/manuscript.tex

Missing data problems are ubiquitous and pervasive. \textbf{or}
Everybody who works with person data is likely to run into a missing
data problem.\\
--\textgreater{} Missingness is problematic because statistical
inference cannot be performed on the whole data without employing
\emph{ad hoc} solutions (e.g., list-wise deletion), which may yield
wildly invalid results \citep{buur18}. A popular method to obtain
statistically valid results despite of missing data is `multiple
imputation' \citep[MI;][]{rubin76}. The multiple imputation method can
be roughly divided into 3 steps: First, each missing value in an
incomplete dataset is `imputed' (i.e., filled in) several times by some
algorithmic procedure. Next, the analysis of scientific interest is
performed on each of the completed datasets. And finally, the results of
these analyses are pooled into a single aggregated statistic.

@@

The complete data is `amputed' once for each simulation repetition with
function
\texttt{mice::ampute().\ (**change\ this:**\ The\ missingness\ is\ univariate,\ and\ the\ probability\ to\ be\ missing\ is\ the\ same\ for\ all\ four\ variables,\ namely\ 20\textbackslash{}\%\ (}prop
= 0.8, mech = ``MCAR''`). This leaves 20\% of the rows completely
observed). \textbf{(Add different percentages missingness??)} This study
only considers only an MCAR missingness mechanism. Therefore, results
may not be extrapolated to other missing data problems. Proper
performance of the convergence diagnostics under MCAR is necessary but
not sufficient to demonstrate appropriateness of \(\widehat{R}\) and
\(AC\) as convergence diagnostics. This is just a proof of concept.

Missing datapoints are imputed with the function \texttt{mice::mice()}.
All MI procedures are performed with Bayesian linear regression
imputation (\texttt{method\ =\ "norm"}), and five imputation chains
(\texttt{m\ =\ 5}). The number of iterations varies between simulation
conditions (\texttt{maxit\ =} \[1, 2, \dots, 100\]).

What to evaluate?

\begin{itemize}
\item
  bias in regression coefficients
\item
  bias in \(R^2\) (= coefficient of determination)
\item
\end{itemize}

>>>>

Missing data problems are ubiquitous and pervasive. \textbf{or}
Everybody who works with person data is likely to run into a missing
data problem.\\
Missingness is problematic because statistical inference cannot be
performed on the whole data without employing \emph{ad hoc} solutions
(e.g., list-wise deletion), which may yield wildly invalid results
\citep{buur18}. A popular method to obtain statistically valid results
despite of missing data is `multiple imputation' \citep[MI;][]{rubin76}.
The multiple imputation method can be roughly divided into 3 steps:
First, each missing value in an incomplete dataset is `imputed' (i.e.,
filled in) several times by some algorithmic procedure. Next, the
analysis of scientific interest is performed on each of the completed
datasets. And finally, the results of these analyses are pooled into a
single aggregated statistic.

@@

The complete data is `amputed' once for each simulation repetition with
function \texttt{mice::ampute()}. (\textbf{change this:} The missingness
is univariate, and the probability to be missing is the same for all
four variables, namely 20\% (\texttt{prop\ =\ 0.8,\ mech\ =\ "MCAR"}).
This leaves 20\% of the rows completely observed). \textbf{(Add
different percentages missingness??)} This study only considers only an
MCAR missingness mechanism. Therefore, results may not be extrapolated
to other missing data problems. Proper performance of the convergence
diagnostics under MCAR is necessary but not sufficient to demonstrate
appropriateness of \(\widehat{R}\) and \(AC\) as convergence
diagnostics. This is just a proof of concept.

Missing datapoints are imputed with the function \texttt{mice::mice()}.
All MI procedures are performed with Bayesian linear regression
imputation (\texttt{method\ =\ "norm"}), and five imputation chains
(\texttt{m\ =\ 5}). The number of iterations varies between simulation
conditions (\texttt{maxit\ =} \(1, 2, \dots, 100\)).
